Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media

### 2015-10-05 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
Updates in the Abstract and first paragraphs of Introd.

\newcommand{\JGnote}[1]{\fbox{\parbox{\textwidth}{ \color{blue} JG Note $\Rightarrow$ #1}}}
\journal{Advances in Water Resources}

@@

\textcolor{blue}{Accurate simulation of multiphase flows in porous media remains a challenge. While the governing equations are the conservation of mass and Darcy's law, there are also the primary variables, such us pressure, velocity, saturation and permeability. This paper introduce a multi-fluid porous media flow model which is the basis of a reservoir simulator, as implemented in Fluidity, a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to control the local solution errors and  increase the computational accuracy in the research on the development of numerical methods in geophysical fluid mechanics. The governing multiphase porous media flow equations are solved using the adaptive and unstructured mesh.
The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
%The element type described above has a discontinuous $n^{th}$ order polynomial representation of velocity and a continuous $(n+1)^{th}$ representation for pressure. This method includes the Darcy force balance equation, pressure, velocity and saturation fields throughout the domain while there is automatic mesh adaptivity in space and time and the computational cost is reduced.} 
Numerical modelling of multiphase flows in porous media has application in a wide range of disciplines, from hydrocarbons exploration and extraction to groundwater operations.Viscous fingering, member of the flow instabilities family, is a major concern for the oil \& gas industry, but not only to this. The impact of fingers formation in our simulations is also investigated.  
%This paper describes a  In particular, the \PNDG[1]{1} element (i.e., discontinuous and piecewise linear shape functions to discretise velocity and pressure fields) is introduced as the basis of the discretisation and also as the CVFEM counterpart. In addition to the unstructured grid used for the simulations, uniform and random permeability distribution are introduced. The formulation is applied to saturated porous media systems.
We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage projects. }

@@

\clearpage
%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Introduction}\label{section:Intro}
Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. This is mainly due to applications on oil and gas recovery, groundwater exploration, underground coal gasification, and more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011} as well as in groundwater operations.

For reservoir engineers and hydrogeologists, mapping the behaviour of multiphase flows and predict the response of the system is a major challenge followed by complex geometries with irregular boundaries and heterogeneous geological formations. These are some of the main challenges for accurately forecasting the behaviour of flows into porous media. Both Permeability and porosity are important parameters in the equations used for numerical simulation of multiphase flow in porous media. %Heterogeniety of permeability and porosity is a product of the medium's formation process %-- deposition and diagenesis -- and exists at different scales. 
At the microscopic level, the continuum approach can be applied to a representative elementary Volume (REV) to define these parameters~\citep{bastian_phd1999}. This is well suited for analysis of small scales such as core samples.

Although parallel processing is available nowadays, a fine-scale geological model is far too detailed to be accumulated by common reservoir simulators, mainly because of the processing time.

@@

The first approach is limited to low flow rates and very small heterogeneity. The second approach overcomes these limitation but also posses drawbacks which include; availability of fine-grid simulation with suitable level of accuracy, dependence of the psuedo functions on rate, direction and grid specification. The experimental data from the third approach is only suitable for a small class of scenarios.~\citep{Chang_1994}. While  non-uniform permeability and porosity may be easy to deal with at a microscopic scale for porous media flow equations, it presents a lot of challenges at a meso/field scale, which is required for reservoir simulations.

@@

\begin{figure}[h]
\makebox[\textwidth]{%
\includegraphics[width=0.49\textwidth]{2b2_P1DGP2_perm}%
\hfill    
\includegraphics[width=0.49\textwidth]{2b2_P1DGP2_mesh}%
}\\[0.5cm]% If you want some vertical space
\makebox[\textwidth]{%
\includegraphics[width=0.75\textwidth]{2b2_P1DGP2_plot}%
}%
\caption{ Caption 1}
\label{fig:1}
\end{figure}

@@

\begin{center}
\centering
\includegraphics[width=0.75\linewidth, height=5cm]{P1DG-P2_element_type} 
\caption{Caption 2}
\label{fig:2}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{4r_po_adapt_fine_125_mesh} 
\caption{Caption 3}
\label{fig:3}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{4r_po_adapt_fine_100_mesh}
\caption{Caption 4}
\label{fig:4}
\end{center}

@@

\begin{landscape}
\begin{figure}[ht!]
\begin{center}
\subfigure[MR100, at the beginning of the simulations]{%
\label{fig:1st}
\includegraphics[width=0.75\textwidth]{4r_po_adapt_fine_100}
}%
\subfigure[MR100, at the middle of the simulation]{%
\label{fig:2nd}
\includegraphics[width=0.75\textwidth]{4r_po_adapt_fine_125_phase1_vol_fraction_mesh}
}\\
\subfigure[MR100, Phase 1 Volume fraction]{%
\label{fig:3rd}
\includegraphics[width=0.75\textwidth]{4r_po_adapt_fine_125_phase1_vol_frac}
}%
\subfigure[MR100, Phase 2 Volume fraction]{%
\label{fig:4th}
\includegraphics[width=0.75\textwidth]{4r_po_adapt_fine_125_phase2_vol_fraction}
}\\
\caption{Comparison of different MR's and their recovery rates}
\label{fig:5}
\end{center}
\end{figure}
\end{landscape}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[MR1]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{MR1.png}
}%
\subfigure[Recovery for MR1]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{Recovery_MR1_1.png}
}\\
\subfigure[MR10]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{MR10.png}
}%
\subfigure[Recovery for MR10]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{Recovery_MR10_10.png}
}\\
\subfigure[MR100]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{MR100_100_start.png}
}%
\subfigure[Recovery for MR100]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{Recovery_MR100_100.png}
}\\ 
\end{center}
\caption{Comparison of different MR's and their recovery rates}
\label{fig:6}
\end{figure}

>>>>

\newcommand{\JGnote}[1]{\fbox{\parbox{\textwidth}{ \color{blue} JG Note $\Rightarrow$ #1}}}
\newcommand{\red}{\textcolor{red}}
\journal{Advances in Water Resources}

@@

Accurate simulation of multiphase flows in porous media remains a challenge \red{(Why? Sentencing is hanging ... word `accurate' is vague here! The second sentence is not gramatically correct and too simplistic -- see notes below.)}. While the governing equations are the conservation of mass and Darcy's law, there are also the primary variables, such us pressure, velocity, saturation and permeability. This paper introduces a multi-fluid porous media flow model which is the basis of a reservoir simulator, as implemented in Fluidity, a general purpose flow simulator that is equipped with anisotropic mesh adaptivity techniques to control the local solution errors and  increase the computational accuracy in the research on the development of numerical methods in geophysical fluid mechanics. The governing multiphase porous media flow equations are solved using the adaptive and unstructured mesh.
The model is based upon two key numerical characteristics: (a) the family of $P_{n}DG-P_{n+1}$ finite element pairs and (b) a consistent overlapping control volume finite element method (CVFEM) formulation.
%The element type described above has a discontinuous $n^{th}$ order polynomial representation of velocity and a continuous $(n+1)^{th}$ representation for pressure. This method includes the Darcy force balance equation, pressure, velocity and saturation fields throughout the domain while there is automatic mesh adaptivity in space and time and the computational cost is reduced.} 
Numerical modelling of multiphase flows in porous media has application in a wide range of disciplines, from hydrocarbons exploration and extraction to groundwater operations.Viscous fingering, member of the flow instabilities family, is a major concern for the oil \& gas industry, but not only to this. The impact of fingers formation in our simulations is also investigated.  
%This paper describes a  In particular, the \PNDG[1]{1} element (i.e., discontinuous and piecewise linear shape functions to discretise velocity and pressure fields) is introduced as the basis of the discretisation and also as the CVFEM counterpart. In addition to the unstructured grid used for the simulations, uniform and random permeability distribution are introduced. The formulation is applied to saturated porous media systems.
We conclude that understanding the multiphase flow properties will allow us to efficiently facilitate more efficient oil \& gas recovery, efficient contaminant removal and better planning of carbon capture and storage projects.

\JGnote{ Abstracts for journal papers have very `squared-structure': (a) Motivation (why would anyone bother to read the paper? Why is it (the paper and/or the subject) relevant? This needs to be covered in 1-3 punchy sentences.); (b) (Main) Objective(s) of the manuscript with particular focus on the \underline{novelty(ies)} the paper may bring (this should be covered in 2-6 (very dry) sentences); (c) Summary of the main technologies that you are using or introducing (and, in some cases, how this/these link(s) with (a), 2-6 sentences); (d) Summary of the main findings (linkage with (b), 2-3 sentences).}

@@

\clearpage
%%%         %%%
%%% SECTION %%%
%%%         %%%
\section{Introduction}\label{section:Intro}
Numerical investigation and modelling of multiphase flows in porous media have attracted the attention of the scientific community over the past 40 years. This is mainly due to applications on oil and gas recovery, groundwater exploration, underground coal gasification, and more recently on CO$_{\text{2}}$ migration and trapping mechanisms in carbon capture and storage (CCS) operations~\citep{spycher_2003, chen_2006, self_2012, pruess_1990c, white_1981,jiang_2011}.

For reservoir engineers and hydrogeologists, mapping \red{(you don't map a behaviour)} the behaviour of multiphase flows and predict the response of the system are major challenges followed by complex geometries with irregular boundaries and heterogeneous geological formations \red{(sentence needs to be improved)}. These are some of the main challenges \red{(too many challenges in the same paragraph ...)} for accurately forecasting the behaviour of flows into porous media. Both permeability and porosity are important parameters in the equations used for numerical simulation of multiphase flow in porous media.\red{(what was the objective of this sentence?)}  %Heterogeniety of permeability and porosity is a product of the medium's formation process %-- deposition and diagenesis -- and exists at different scales. 
At the microscopic level, the continuum approach can be applied to a representative elementary Volume (REV) to define these parameters~\citep{bastian_phd1999}. This is well suited for analysis of small scales such as core samples \red{(These sentences are not linked with the previous sentences, also what is the objective of them? )}.

Although parallel processing is available nowadays, a fine-scale geological model is far too detailed to be accumulated by common reservoir simulators, mainly because of the processing time. \red{(Same comment as above!)}

@@

The first approach is limited to low flow rates and very small heterogeneity. The second approach overcomes these limitation but also posses drawbacks which include; availability of fine-grid simulation with suitable level of accuracy, dependence of the psuedo functions on rate, direction and grid specification. The experimental data from the third approach is only suitable for a small class of scenarios.~\citep{Chang_1994}. While  non-uniform permeability and porosity may be easy to deal with at a microscopic scale for porous media flow equations, it presents a lot of challenges at a meso/field scale, which is required for reservoir simulations. \red{From `At the microscopic level ...' to `...reservoir simulations' --try to rewrite these sentences/paragraphs ... they are very confusing and not really connected'}

\JGnote{As a rule of thumb, when you are introducing a paragraph, the very first sentence needs to give the reader a clue of either linkage w previous paragraphs or linkage w the main subject(s).}

@@

\begin{figure}[h]
\makebox[\textwidth]{%
\includegraphics[width=0.49\textwidth]{./Pics/2b2_P1DGP2_perm}%
\hfill    
\includegraphics[width=0.49\textwidth]{./Pics/2b2_P1DGP2_mesh}%
}\\[0.5cm]% If you want some vertical space
\makebox[\textwidth]{%
\includegraphics[width=0.75\textwidth]{./Pics/2b2_P1DGP2_plot}%
}%
\caption{ Caption 1}
\label{fig:1}
\end{figure}

@@

\begin{center}
\centering
\includegraphics[width=0.75\linewidth, height=5cm]{./Pics/P1DG-P2_element_type} 
\caption{Caption 2}
\label{fig:2}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{./Pics/4r_po_adapt_fine_125_mesh} 
\caption{Caption 3}
\label{fig:3}
\end{center}

\begin{center}
\centering
\includegraphics[width=0.8\linewidth, height=7cm]{./Pics/4r_po_adapt_fine_100_mesh}
\caption{Caption 4}
\label{fig:4}
\end{center}

@@

\begin{landscape}
\begin{figure}[ht!]
\begin{center}
\subfigure[MR100, at the beginning of the simulations]{%
\label{fig:1st}
\includegraphics[width=0.75\textwidth]{./Pics/4r_po_adapt_fine_100}
}%
\subfigure[MR100, at the middle of the simulation]{%
\label{fig:2nd}
\includegraphics[width=0.75\textwidth]{./Pics/4r_po_adapt_fine_125_phase1_vol_fraction_mesh}
}\\
\subfigure[MR100, Phase 1 Volume fraction]{%
\label{fig:3rd}
\includegraphics[width=0.75\textwidth]{./Pics/4r_po_adapt_fine_125_phase1_vol_frac}
}%
\subfigure[MR100, Phase 2 Volume fraction]{%
\label{fig:4th}
\includegraphics[width=0.75\textwidth]{./Pics/4r_po_adapt_fine_125_phase2_vol_fraction}
}\\
\caption{Comparison of different MR's and their recovery rates}
\label{fig:5}
\end{center}
\end{figure}
\end{landscape}

@@

\begin{figure}[ht!]
\begin{center}
\subfigure[MR1]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{./Pics/MR1.png}
}%
\subfigure[Recovery for MR1]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR1_1.png}
}\\
\subfigure[MR10]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{./Pics/MR10.png}
}%
\subfigure[Recovery for MR10]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR10_10.png}
}\\
\subfigure[MR100]{%
\label{fig:first}
\includegraphics[width=0.7\textwidth]{./Pics/MR100_100_start.png}
}%
\subfigure[Recovery for MR100]{%
\label{fig:second}
\includegraphics[width=0.52\textwidth]{./Pics/Recovery_MR100_100.png}
}\\ 
\end{center}
\caption{Comparison of different MR's and their recovery rates}
\label{fig:6}
\end{figure}
